Manuscript results: Forecasting influenza using ARIMA and ARIMAX

### 2025-04-25 manuscript_results.Rmd
update

---
title: "Final manuscript"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
---

This document summarizes the results of the 12 models that were developed: AUTO ARIMA, AUTO ADJACENT, AUTO EPIWEEK, ES27 TEMPERATURE, ES27 ARIMA, ES27 ADJACENT, ES27 EPIWEEK, ES64 TEMPERATURE, ES64 ARIMA, ES64 ADJACENT, ES64 EPIWEEK, ES64 TEMPERATURE. It has a series of histograms and maps that compare the results of these models based on their WIS. You can read more about the specific covariates and ensemble techniques of these models in their own Rmarkdown documents.

@@

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("dplyr")
library("scales")  # For `breaks_extended`
library("ggplot2")
library("broom")
library("fields")

@@

```{r}
load("NHSH_hospitalizations_seed_nolog_correct/ES_ARIMA/ARIMA_MODELS_influenza_hospitalization_nolog.Rdata")
load("NHSH_hospitalizations_seed_nolog_correct/ES_ADJACENT/ADJACENT_MODELS_influenza_hospitalization_nolog.Rdata")
load("NHSH_hospitalizations_seed_nolog_correct/ES_EPIWEEK/EPIWEEK_MODELS_influenza_hospitalization_nolog.Rdata")
load("NHSH_hospitalizations_seed_nolog_correct/ES_TEMPERATURE/TEMPERATURE_MODELS_influenza_hospitalization_nolog.Rdata")
load("NHSH_hospitalizations_seed_nolog_correct/ES_AVERAGE/AVERAGE_MODELS_influenza_hospitalization_nolog.Rdata")

@@

# Dataframe that will be analysed for 1 Week Ahead
df_W1_NoLg <- data.frame(
  STATE = AUTO_ARIMA_WEEK1$State,
  Julian_date = AUTO_ARIMA_WEEK1$target_end_date,
  epiweek = AUTO_ARIMA_WEEK1$epiweek,
  AUTO_AR_NoLg=AUTO_ARIMA_WEEK1$WIS, 
  AUTO_ADJ_NoLg=AUTO_ADJACENT_WEEK1$WIS, 
  AUTO_EPI_NoLg=AUTO_EPIWEEK_WEEK1$WIS,
  AUTO_TEMP_NoLg=AUTO_TEMPERATURE_WEEK1$WIS,
  AUTO_AVG_NoLg=AUTO_AVERAGE_WEEK1$WIS,
  ES27_AR_NoLg=ES27_ARIMA_WEEK1$WIS, 
  ES27_ADJ_NoLg=ES27_ADJACENT_WEEK1$WIS, 
  ES27_EPI_NoLg=ES27_EPIWEEK_WEEK1$WIS,
  ES27_TEMP_NoLg=ES27_TEMPERATURE_WEEK1$WIS,
  ES27_AVG_NoLg=ES27_AVERAGE_WEEK1$WIS,
  ES64_AR_NoLg=ES64_ARIMA_WEEK1$WIS, 
  ES64_ADJ_NoLg=ES64_ADJACENT_WEEK1$WIS, 
  ES64_EPI_NoLg=ES64_EPIWEEK_WEEK1$WIS,
  ES64_TEMP_NoLg=ES64_TEMPERATURE_WEEK1$WIS,
  ES64_AVG_NoLg=ES64_AVERAGE_WEEK1$WIS
)

@@

@@

# Combine all datasets and add a column to identify them
df_combined_ <- bind_rows(
  df_W1_NoLg %>% mutate(Dataset = "W1"),
  df_W2_NoLg %>% mutate(Dataset = "W2"),
  df_W3_NoLg %>% mutate(Dataset = "W3"),
  df_W4_NoLg %>% mutate(Dataset = "W4")
) %>%
  select(STATE, epiweek, AUTO_AR_NoLg, Dataset, Julian_date) %>%
  rename(WIS = AUTO_AR_NoLg)  # Rename for consistency

@@

@@

# Define the function
calculate_mean_wis <- function(data) {
  data %>%
    group_by(STATE) %>%
    summarize(
      AUTO_AR_NoLg = mean(AUTO_AR_NoLg, na.rm = TRUE),
      AUTO_ADJ_NoLg = mean(AUTO_ADJ_NoLg, na.rm = TRUE),
      AUTO_EPI_NoLg = mean(AUTO_EPI_NoLg, na.rm = TRUE),
      AUTO_TMP_NoLg = mean(AUTO_TEMP_NoLg, na.rm = TRUE),
      AUTO_AVG_NoLg = mean(AUTO_AVG_NoLg, na.rm = TRUE),
      ES27_AR_NoLg = mean(ES27_AR_NoLg, na.rm = TRUE),
      ES27_ADJ_NoLg = mean(ES27_ADJ_NoLg, na.rm = TRUE),
      ES27_EPI_NoLg = mean(ES27_EPI_NoLg, na.rm = TRUE),
      ES27_TMP_NoLg = mean(ES27_TEMP_NoLg, na.rm = TRUE),
      ES27_AVG_NoLg = mean(ES27_AVG_NoLg, na.rm = TRUE),
      ES64_AR_NoLg = mean(ES64_AR_NoLg, na.rm = TRUE),
      ES64_ADJ_NoLg = mean(ES64_ADJ_NoLg, na.rm = TRUE),
      ES64_EPI_NoLg = mean(ES64_EPI_NoLg, na.rm = TRUE),
      ES64_TMP_NoLg = mean(ES64_TEMP_NoLg, na.rm = TRUE),
      ES64_AVG_NoLg = mean(ES64_AVG_NoLg, na.rm = TRUE)

@@

@@

# Define color ramps for each category
colors_ar <- c("#a6cee3","#226e83", "#08306b")
colors_adj <- c("#fb9a99","red3", "#a11c3e")
colors_epi <- c("#cab2d6", "purple2", "#5e2b7b")
colors_tmp <- c("lightgreen","green3","#319045")

# Create a custom color mapping for each variable
custom_colors <- c(
  "AUTO_AR" = colors_ar[1], "ES27_AR" = colors_ar[2], "ES64_AR" = colors_ar[3],
  "AUTO_EPI" = colors_epi[1], "ES27_EPI" = colors_epi[2], "ES64_EPI" = colors_epi[3],
  "AUTO_TMP" = colors_tmp[1], "ES27_TMP" = colors_tmp[2], "ES64_TMP" = colors_tmp[3],
  "AUTO_ADJ" = colors_adj[1], "ES27_ADJ" = colors_adj[2], "ES64_ADJ" = colors_adj[3],
  "AUTO_AVG" = colors_adj[1], "ES27_AVG" = colors_adj[2], "ES64_AVG" = colors_adj[3]
)

  # Create the box plot with additional facet for categories
plt1<-  ggplot(combined_data, aes(x = week, y = value, fill = Models)) +
    geom_boxplot(color = "black") +
    scale_fill_manual(values = custom_colors) +
    facet_wrap(category ~ Models , nrow = 2) +
    labs(
      title = "A) Models without log-back transformations",
      x = "",
      y = "Mean Weighted Interval Score (WIS)"
    ) +
  scale_y_continuous(limits = c(0, 650)) +  # Set y-axis range
    theme_minimal()
plt1
ggsave(plt1, height = 8, width = 10, filename="fig1A.jpg")  
```

```{r}
load("NHSH_hospitalizations_seed/ES_ARIMA/ARIMA_MODELS_influenza_hospitalization.Rdata")
load("NHSH_hospitalizations_seed/ES_ADJACENT/ADJACENT_MODELS_influenza_hospitalization.Rdata")
load("NHSH_hospitalizations_seed/ES_EPIWEEK/EPIWEEK_MODELS_influenza_hospitalization.Rdata")
load("NHSH_hospitalizations_seed/ES_TEMPERATURE/TEMPERATURE_MODELS_influenza_hospitalization.Rdata")
load("NHSH_hospitalizations_seed/ES_AVERAGE/AVERAGE_MODELS_influenza_hospitalization.Rdata")

states <- read_sf("NHSH_hospitalizations_seed/shapefiles/cb_2018_us_state_500k.shp")

@@

# Define the function
calculate_mean_wis <- function(data) {
  data %>%
    group_by(STATE) %>%
    summarize(
      AUTO_AR_Lg = mean(AUTO_AR, na.rm = TRUE),
      AUTO_ADJ_Lg = mean(AUTO_ADJ, na.rm = TRUE),
      AUTO_EPI_Lg = mean(AUTO_EPI, na.rm = TRUE),
      AUTO_TMP_Lg = mean(AUTO_TEMP, na.rm = TRUE),
      ES27_AR_Lg = mean(ES27_AR, na.rm = TRUE),
      ES27_ADJ_Lg = mean(ES27_ADJ, na.rm = TRUE),
      ES27_EPI_Lg = mean(ES27_EPI, na.rm = TRUE),
      ES27_TMP_Lg = mean(ES27_TEMP, na.rm = TRUE),
      ES64_AR_Lg = mean(ES64_AR, na.rm = TRUE),
      ES64_ADJ_Lg = mean(ES64_ADJ, na.rm = TRUE),
      ES64_EPI_Lg = mean(ES64_EPI, na.rm = TRUE),
      ES64_TMP_Lg = mean(ES64_TEMP, na.rm = TRUE)
    )
}

@@

  Models = rep(c("AUTO_AR_Lg", "ES27_AR_Lg", "ES64_AR_Lg", "AUTO_ADJ_Lg", "ES27_ADJ_Lg", "ES64_ADJ_Lg", "AUTO_TMP_Lg", "ES27_TMP_Lg", "ES64_TMP_Lg", "AUTO_EPI_Lg", "ES27_EPI_Lg", "ES64_EPI_Lg"), each=48, times = 4),

  value = c(W1_Lg$AUTO_AR_Lg, W1_Lg$ES27_AR_Lg, W1_Lg$ES64_AR_Lg, W1_Lg$AUTO_ADJ_Lg, W1_Lg$ES27_ADJ_Lg, W1_Lg$ES64_ADJ_Lg, W1_Lg$AUTO_TMP_Lg, W1_Lg$ES27_TMP_Lg, W1_Lg$ES64_TMP_Lg, W1_Lg$AUTO_EPI_Lg, W1_Lg$ES27_EPI_Lg, W1_Lg$ES64_EPI_Lg,

            W2_Lg$AUTO_AR_Lg, W2_Lg$ES27_AR_Lg, W2_Lg$ES64_AR_Lg, W2_Lg$AUTO_ADJ_Lg, W2_Lg$ES27_ADJ_Lg, W2_Lg$ES64_ADJ_Lg, W2_Lg$AUTO_TMP_Lg, W2_Lg$ES27_TMP_Lg, W2_Lg$ES64_TMP_Lg, W2_Lg$AUTO_EPI_Lg, W2_Lg$ES27_EPI_Lg, W2_Lg$ES64_EPI_Lg,

            W3_Lg$AUTO_AR_Lg, W3_Lg$ES27_AR_Lg, W3_Lg$ES64_AR_Lg, W3_Lg$AUTO_ADJ_Lg, W3_Lg$ES27_ADJ_Lg, W3_Lg$ES64_ADJ_Lg, W3_Lg$AUTO_TMP_Lg, W3_Lg$ES27_TMP_Lg, W3_Lg$ES64_TMP_Lg, W3_Lg$AUTO_EPI_Lg, W3_Lg$ES27_EPI_Lg, W3_Lg$ES64_EPI_Lg,

            W4_Lg$AUTO_AR_Lg, W4_Lg$ES27_AR_Lg, W4_Lg$ES64_AR_Lg, W4_Lg$AUTO_ADJ_Lg, W4_Lg$ES27_ADJ_Lg, W4_Lg$ES64_ADJ_Lg, W4_Lg$AUTO_TMP_Lg, W4_Lg$ES27_TMP_Lg, W4_Lg$ES64_TMP_Lg, W4_Lg$AUTO_EPI_Lg, W4_Lg$ES27_EPI_Lg, W4_Lg$ES64_EPI_Lg)
)

# Create a new column for categories
combined_data <- combined_data %>%
  mutate(category = case_when(
    grepl("AR", Models) ~ "AR",
    grepl("EPI", Models) ~ "EPI",
    grepl("TMP", Models) ~ "TMP",
    grepl("ADJ", Models) ~ "ADJ"
  ))

# Define color ramps for each category
colors_ar <- c("#a6cee3","#226e83", "#08306b")
colors_adj <- c("#fb9a99","red3", "#a11c3e")
colors_epi <- c("#cab2d6", "purple2", "#5e2b7b")
colors_tmp <- c("lightgreen","green3","#319045")

# Create a custom color mapping for each variable
custom_colors <- c(
  "AUTO_AR_Lg" = colors_ar[1], "ES27_AR_Lg" = colors_ar[2], "ES64_AR_Lg" = colors_ar[3],
  "AUTO_EPI_Lg" = colors_epi[1], "ES27_EPI_Lg" = colors_epi[2], "ES64_EPI_Lg" = colors_epi[3],
  "AUTO_TMP_Lg" = colors_tmp[1], "ES27_TMP_Lg" = colors_tmp[2], "ES64_TMP_Lg" = colors_tmp[3],
  "AUTO_ADJ_Lg" = colors_adj[1], "ES27_ADJ_Lg" = colors_adj[2], "ES64_ADJ_Lg" = colors_adj[3]
)

  # Create the box plot with additional facet for categories
plt2<-  ggplot(combined_data, aes(x = week, y = value, fill = Models)) +
    geom_boxplot(color = "black") +
    scale_fill_manual(values = custom_colors) +
    facet_wrap(category ~ Models , nrow = 2) +
    labs(
      title = "B)Models with log-back transformations",
      x = "",
      y = "Weighted Interval Score (WIS)"
    ) +
  scale_y_continuous(limits = c(0, 650)) +  # Set y-axis range
    theme_minimal()
plt2
ggsave(plt2, height = 8, width = 10, filename="Fig1B.jpg")  
```

@@

```{r}
W1<-merge(W1_Lg,W1_NoLg, by = "STATE")
W2<-merge(W2_Lg,W2_NoLg, by = "STATE")
W3<-merge(W3_Lg,W3_NoLg, by = "STATE")
W4<-merge(W4_Lg,W4_NoLg, by = "STATE")

@@

# Create a custom color mapping for each variable
custom_colors <- c(
  "AUTO_AR_Lg" = colors_ar[1], "ES27_AR_Lg" = colors_ar[2], "ES64_AR_Lg" = colors_ar[3],
  "AUTO_EPI_Lg" = colors_epi[1], "ES27_EPI_Lg" = colors_epi[2], "ES64_EPI_Lg" = colors_epi[3],
  "AUTO_TMP_Lg" = colors_tmp[1], "ES27_TMP_Lg" = colors_tmp[2], "ES64_TMP_Lg" = colors_tmp[3],
  "AUTO_ADJ_Lg" = colors_adj[1], "ES27_ADJ_Lg" = colors_adj[2], "ES64_ADJ_Lg" = colors_adj[3],

  "AUTO_AR_NoLg" = colors_ar[1], "ES27_AR_NoLg" = colors_ar[2], "ES64_AR_NoLg" = colors_ar[3],
  "AUTO_EPI_NoLg" = colors_epi[1], "ES27_EPI_NoLg" = colors_epi[2], "ES64_EPI_NoLg" = colors_epi[3],
  "AUTO_TMP_NoLg" = colors_tmp[1], "ES27_TMP_NoLg" = colors_tmp[2], "ES64_TMP_NoLg" = colors_tmp[3],
  "AUTO_ADJ_NoLg" = colors_adj[1], "ES27_ADJ_NoLg" = colors_adj[2], "ES64_ADJ_NoLg" = colors_adj[3]
)

@@

plot2
# --------- WEEK3 MODELS ------------- #
plot3<-ggplot(W3, aes(x = Best_Result, fill = Best_Result))  +  
  geom_bar() +  
  scale_fill_manual(values = custom_colors) +  
  labs(title = "Best model on the 48 U.S. states (W3)",       
       x = "", y = "Number of states") +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  facet_wrap(~ Model_Type, scales = "free_x")
plot3

@@

 Wilcox test with Holm p-value ajustment

1 WEEK AHEAD

@@

2 WEEKS AHEAD Post Hoc Wilcox test

@@

>>>>

---
title: "Manuscript results: Forecasting influenza using ARIMA and ARIMAX"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
---

This document has the results that we present in our manuscript draft.

@@

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("dplyr")
library("scales")  # For `breaks_extended`
library("ggplot2")
library("broom")
library("fields")
library("ggpubr")
```

@@

load("NHSH_hospitalizations_seed_nolog_correct/ES_ARIMA/ARIMA_MODELS_influenza_hospitalization_nolog.Rdata")
load("NHSH_hospitalizations_seed_nolog_correct/ES_ADJACENT/ADJACENT_MODELS_influenza_hospitalization_nolog.Rdata")
load("NHSH_hospitalizations_seed_nolog_correct/ES_EPIWEEK/EPIWEEK_MODELS_influenza_hospitalization_nolog.Rdata")
load("NHSH_hospitalizations_seed_nolog_correct/ES_TEMPERATURE/TEMPERATURE_MODELS_influenza_hospitalization_nolog.Rdata")
load("NHSH_hospitalizations_seed_nolog_correct/ES_AVERAGE/AVERAGE_MODELS_influenza_hospitalization_nolog.Rdata")

@@

# Dataframe that will be analysed for 1 Week Ahead
df_W1_NoLg <- data.frame(
  STATE = AUTO_ARIMA_WEEK1$State,
  Julian_date = AUTO_ARIMA_WEEK1$target_end_date,
  epiweek = AUTO_ARIMA_WEEK1$epiweek,
  AUTO_AR_NoLg=AUTO_ARIMA_WEEK1$WIS, 
  AUTO_ADJ_NoLg=AUTO_ADJACENT_WEEK1$WIS, 
  AUTO_EPI_NoLg=AUTO_EPIWEEK_WEEK1$WIS,
  AUTO_TEMP_NoLg=AUTO_TEMPERATURE_WEEK1$WIS,
  AUTO_AVG_NoLg=AUTO_AVERAGE_WEEK1$WIS,

  ES27_AR_NoLg=ES27_ARIMA_WEEK1$WIS, 
  ES27_ADJ_NoLg=ES27_ADJACENT_WEEK1$WIS, 
  ES27_EPI_NoLg=ES27_EPIWEEK_WEEK1$WIS,
  ES27_TEMP_NoLg=ES27_TEMPERATURE_WEEK1$WIS,
  ES27_AVG_NoLg=ES27_AVERAGE_WEEK1$WIS,

  ES64_AR_NoLg=ES64_ARIMA_WEEK1$WIS, 
  ES64_ADJ_NoLg=ES64_ADJACENT_WEEK1$WIS, 
  ES64_EPI_NoLg=ES64_EPIWEEK_WEEK1$WIS,
  ES64_TEMP_NoLg=ES64_TEMPERATURE_WEEK1$WIS,
  ES64_AVG_NoLg=ES64_AVERAGE_WEEK1$WIS
)

@@

Example of influenza hospitalizations forecasts and 95% confidence interval.

# Create a data frame with forecasts for plotting
forecast_df <- data.frame(
  Week = factor(1:4, labels = paste0("Week ", 1:4)),  # Label weeks
  Value = c(AUTO_ARIMA_WEEK1_list[[1]]$forecasts[5], 
            AUTO_ARIMA_WEEK2_list[[1]]$forecasts[5], 
            AUTO_ARIMA_WEEK3_list[[1]]$forecasts[5], 
            AUTO_ARIMA_WEEK4_list[[1]]$forecasts[5]),
  Lower = c(AUTO_ARIMA_WEEK1_list[[1]][["0.025"]][5], 
            AUTO_ARIMA_WEEK2_list[[1]][["0.025"]][5], 
            AUTO_ARIMA_WEEK3_list[[1]][["0.025"]][5], 
            AUTO_ARIMA_WEEK4_list[[1]][["0.025"]][5]),
  Upper = c(AUTO_ARIMA_WEEK1_list[[1]][["0.975"]][5], 
            AUTO_ARIMA_WEEK2_list[[1]][["0.975"]][5], 
            AUTO_ARIMA_WEEK3_list[[1]][["0.975"]][5], 
            AUTO_ARIMA_WEEK4_list[[1]][["0.975"]][5])
)

# Plotting forecasts and ribbon
plt00 <- ggplot(forecast_df, aes(x = Week, y = Value, group = 1)) +
  geom_line(color = "black", size = 1) +          # Line connecting forecast values
  geom_point(color = "black", size = 3) +         # Points for each forecast value
  geom_ribbon(aes(ymin = Lower, ymax = Upper),   # Ribbon for confidence intervals
              fill = "gray", alpha = 0.25) +
  labs(title = "Example of influenza hospitalizaions forecasts with 95% confidence intervals",
       x = "Target Weeks",
       y = "Forecast Values") +
  theme_minimal()

plt00

ggsave(plt00, height = 5, width = 6, filename="fig00.jpg") 
```

Weighted interval scores for each state and each target week.

@@

# Combine all datasets and add a column to identify them
df_combined_ <- bind_rows(
  df_W1_NoLg %>% mutate(Dataset = "1 week ahead"),
  df_W2_NoLg %>% mutate(Dataset = "2 weeks ahead"),
  df_W3_NoLg %>% mutate(Dataset = "3 weeks ahead"),
  df_W4_NoLg %>% mutate(Dataset = "4 weeks ahead")
) %>%
  select(STATE, epiweek, AUTO_AR_NoLg, Dataset, Julian_date) %>%
  rename(WIS = AUTO_AR_NoLg)  # Rename for consistency

@@

Computing mean weighted interval score for all target week for each model.

@@

# Define the function
calculate_mean_wis <- function(data) {
  data %>%
    group_by(STATE) %>%
    summarize(
      AUTO_AR_NoLg = mean(AUTO_AR_NoLg, na.rm = TRUE),
      AUTO_ADJ_NoLg = mean(AUTO_ADJ_NoLg, na.rm = TRUE),
      AUTO_EPI_NoLg = mean(AUTO_EPI_NoLg, na.rm = TRUE),
      AUTO_TMP_NoLg = mean(AUTO_TEMP_NoLg, na.rm = TRUE),
      AUTO_AVG_NoLg = mean(AUTO_AVG_NoLg, na.rm = TRUE),

      ES27_AR_NoLg = mean(ES27_AR_NoLg, na.rm = TRUE),
      ES27_ADJ_NoLg = mean(ES27_ADJ_NoLg, na.rm = TRUE),
      ES27_EPI_NoLg = mean(ES27_EPI_NoLg, na.rm = TRUE),
      ES27_TMP_NoLg = mean(ES27_TEMP_NoLg, na.rm = TRUE),
      ES27_AVG_NoLg = mean(ES27_AVG_NoLg, na.rm = TRUE),

      ES64_AR_NoLg = mean(ES64_AR_NoLg, na.rm = TRUE),
      ES64_ADJ_NoLg = mean(ES64_ADJ_NoLg, na.rm = TRUE),
      ES64_EPI_NoLg = mean(ES64_EPI_NoLg, na.rm = TRUE),
      ES64_TMP_NoLg = mean(ES64_TEMP_NoLg, na.rm = TRUE),
      ES64_AVG_NoLg = mean(ES64_AVG_NoLg, na.rm = TRUE)
    )
}

@@

Boxplots of the mean(WIS) by each state.

@@

# Define color ramps for each category
colors_ar <- c("#a6cee3","#226e83", "#08306b")
colors_adj <- c("#fb9a99","red3", "#a11c3e")
colors_epi <- c("#cab2d6", "purple2", "#5e2b7b")
colors_tmp <- c("lightgreen","green3","#319045")
colors_avg <- c("orange","orange2", "orange3")

# Create a custom color mapping for each variable
custom_colors <- c(
  "AUTO_AR" = colors_ar[1], "ES27_AR" = colors_ar[2], "ES64_AR" = colors_ar[3],
  "AUTO_EPI" = colors_epi[1], "ES27_EPI" = colors_epi[2], "ES64_EPI" = colors_epi[3],
  "AUTO_TMP" = colors_tmp[1], "ES27_TMP" = colors_tmp[2], "ES64_TMP" = colors_tmp[3],
  "AUTO_ADJ" = colors_adj[1], "ES27_ADJ" = colors_adj[2], "ES64_ADJ" = colors_adj[3],
  "AUTO_AVG" = colors_avg[1], "ES27_AVG" = colors_avg[2], "ES64_AVG" = colors_avg[3]
)

  # Create the box plot with additional facet for categories
plt1<-  ggplot(combined_data, aes(x = week, y = value, fill = Models)) +
    geom_boxplot(color = "black") +
    scale_fill_manual(values = custom_colors) +
    facet_wrap(category ~ Models , nrow = 3) +
    labs(
      title = "A) Models without log-back transformations",
      x = "Target Week",
      y = "Mean Weighted Interval Score (WIS)"
    ) +
  scale_y_continuous(limits = c(0, 650)) +  # Set y-axis range
    theme_minimal()
plt1
ggsave(plt1, height = 8, width = 10, filename="fig1A.jpg")  
```

Loading model with log-back transformations.

```{r}
load("NHSH_hospitalizations_seed/ES_ARIMA/ARIMA_MODELS_influenza_hospitalization.Rdata")
load("NHSH_hospitalizations_seed/ES_ADJACENT/ADJACENT_MODELS_influenza_hospitalization.Rdata")
load("NHSH_hospitalizations_seed/ES_EPIWEEK/EPIWEEK_MODELS_influenza_hospitalization.Rdata")
load("NHSH_hospitalizations_seed/ES_TEMPERATURE/TEMPERATURE_MODELS_influenza_hospitalization.Rdata")
load("NHSH_hospitalizations_seed/ES_AVERAGE/AVERAGE_MODELS_influenza_hospitalization.Rdata")
```

@@

# Define the function
calculate_mean_wis <- function(data) {
  data %>%
    group_by(STATE) %>%
    summarize(
      AUTO_AR_Lg = mean(AUTO_AR, na.rm = TRUE),
      AUTO_ADJ_Lg = mean(AUTO_ADJ, na.rm = TRUE),
      AUTO_EPI_Lg = mean(AUTO_EPI, na.rm = TRUE),
      AUTO_TMP_Lg = mean(AUTO_TEMP, na.rm = TRUE),
      AUTO_AVG_Lg = mean(AUTO_AVG, na.rm = TRUE),

      ES27_AR_Lg = mean(ES27_AR, na.rm = TRUE),
      ES27_ADJ_Lg = mean(ES27_ADJ, na.rm = TRUE),
      ES27_EPI_Lg = mean(ES27_EPI, na.rm = TRUE),
      ES27_TMP_Lg = mean(ES27_TEMP, na.rm = TRUE),
      ES27_AVG_Lg = mean(ES27_AVG, na.rm = TRUE),

      ES64_AR_Lg = mean(ES64_AR, na.rm = TRUE),
      ES64_ADJ_Lg = mean(ES64_ADJ, na.rm = TRUE),
      ES64_EPI_Lg = mean(ES64_EPI, na.rm = TRUE),
      ES64_TMP_Lg = mean(ES64_TEMP, na.rm = TRUE),
      ES64_AVG_Lg = mean(ES64_AVG, na.rm = TRUE),

@@

  Models = rep(c("AUTO_AR_Lg", "ES27_AR_Lg", "ES64_AR_Lg", "AUTO_ADJ_Lg", "ES27_ADJ_Lg", "ES64_ADJ_Lg", "AUTO_TMP_Lg", "ES27_TMP_Lg", "ES64_TMP_Lg", "AUTO_EPI_Lg", "ES27_EPI_Lg", "ES64_EPI_Lg", "AUTO_AVG_Lg", "ES27_AVG_Lg", "ES64_AVG_Lg"), each=48, times = 4),

  value = c(W1_Lg$AUTO_AR_Lg, W1_Lg$ES27_AR_Lg, W1_Lg$ES64_AR_Lg, 
            W1_Lg$AUTO_ADJ_Lg, W1_Lg$ES27_ADJ_Lg, W1_Lg$ES64_ADJ_Lg, 
            W1_Lg$AUTO_TMP_Lg, W1_Lg$ES27_TMP_Lg, W1_Lg$ES64_TMP_Lg, 
            W1_Lg$AUTO_EPI_Lg, W1_Lg$ES27_EPI_Lg, W1_Lg$ES64_EPI_Lg,
            W1_Lg$AUTO_AVG_Lg, W1_Lg$ES27_AVG_Lg, W1_Lg$ES64_AVG_Lg,

            W2_Lg$AUTO_AR_Lg, W2_Lg$ES27_AR_Lg, W2_Lg$ES64_AR_Lg, 
            W2_Lg$AUTO_ADJ_Lg, W2_Lg$ES27_ADJ_Lg, W2_Lg$ES64_ADJ_Lg, 
            W2_Lg$AUTO_TMP_Lg, W2_Lg$ES27_TMP_Lg, W2_Lg$ES64_TMP_Lg, 
            W2_Lg$AUTO_EPI_Lg, W2_Lg$ES27_EPI_Lg, W2_Lg$ES64_EPI_Lg,
            W2_Lg$AUTO_AVG_Lg, W2_Lg$ES27_AVG_Lg, W2_Lg$ES64_AVG_Lg,

            W3_Lg$AUTO_AR_Lg, W3_Lg$ES27_AR_Lg, W3_Lg$ES64_AR_Lg, 
            W3_Lg$AUTO_ADJ_Lg, W3_Lg$ES27_ADJ_Lg, W3_Lg$ES64_ADJ_Lg, 
            W3_Lg$AUTO_TMP_Lg, W3_Lg$ES27_TMP_Lg, W3_Lg$ES64_TMP_Lg, 
            W3_Lg$AUTO_EPI_Lg, W3_Lg$ES27_EPI_Lg, W3_Lg$ES64_EPI_Lg,
            W3_Lg$AUTO_AVG_Lg, W3_Lg$ES27_AVG_Lg, W3_Lg$ES64_AVG_Lg,

            W4_Lg$AUTO_AR_Lg, W4_Lg$ES27_AR_Lg, W4_Lg$ES64_AR_Lg, 
            W4_Lg$AUTO_ADJ_Lg, W4_Lg$ES27_ADJ_Lg, W4_Lg$ES64_ADJ_Lg, 
            W4_Lg$AUTO_TMP_Lg, W4_Lg$ES27_TMP_Lg, W4_Lg$ES64_TMP_Lg, 
            W4_Lg$AUTO_EPI_Lg, W4_Lg$ES27_EPI_Lg, W4_Lg$ES64_EPI_Lg,
            W4_Lg$AUTO_AVG_Lg, W4_Lg$ES27_AVG_Lg, W4_Lg$ES64_AVG_Lg)

# Define color ramps for each category
colors_ar <- c("#a6cee3","#226e83", "#08306b")
colors_adj <- c("#fb9a99","red3", "#a11c3e")
colors_epi <- c("#cab2d6", "purple2", "#5e2b7b")
colors_tmp <- c("lightgreen","green3","#319045")
colors_avg <- c("orange","orange2", "orange3")

# Create a custom color mapping for each variable
custom_colors <- c(
  "AUTO_AR_Lg" = colors_ar[1], "ES27_AR_Lg" = colors_ar[2], "ES64_AR_Lg" = colors_ar[3],
  "AUTO_EPI_Lg" = colors_epi[1], "ES27_EPI_Lg" = colors_epi[2], "ES64_EPI_Lg" = colors_epi[3],
  "AUTO_TMP_Lg" = colors_tmp[1], "ES27_TMP_Lg" = colors_tmp[2], "ES64_TMP_Lg" = colors_tmp[3],
  "AUTO_ADJ_Lg" = colors_adj[1], "ES27_ADJ_Lg" = colors_adj[2], "ES64_ADJ_Lg" = colors_adj[3],
  "AUTO_AVG_Lg" = colors_avg[1], "ES27_AVG_Lg" = colors_avg[2], "ES64_AVG_Lg" = colors_avg[3]
)

  # Create the box plot with additional facet for categories
plt2<-  ggplot(combined_data, aes(x = week, y = value, fill = Models)) +
    geom_boxplot(color = "black") +
    scale_fill_manual(values = custom_colors) +
    facet_wrap(category ~ Models , nrow = 3) +
    labs(
      title = "B) Models with log-back transformations",
      x = "",
      y = "Weighted Interval Score (WIS)"
    ) +
  scale_y_continuous(limits = c(0, 650)) +  # Set y-axis range
    theme_minimal()
plt2
ggsave(plt2, height = 8, width = 10, filename="Fig1B.jpg")  
```

@@

```{r}
W1<-merge(W1_Lg,W1_NoLg, by = "STATE")
W2<-merge(W2_Lg,W2_NoLg, by = "STATE")
W3<-merge(W3_Lg,W3_NoLg, by = "STATE")
W4<-merge(W4_Lg,W4_NoLg, by = "STATE")
```

@@

# Create a custom color mapping for each variable
custom_colors <- c(
  "AUTO_AR_Lg" = colors_ar[1], "ES27_AR_Lg" = colors_ar[2], "ES64_AR_Lg" = colors_ar[3],
  "AUTO_EPI_Lg" = colors_epi[1], "ES27_EPI_Lg" = colors_epi[2], "ES64_EPI_Lg" = colors_epi[3],
  "AUTO_TMP_Lg" = colors_tmp[1], "ES27_TMP_Lg" = colors_tmp[2], "ES64_TMP_Lg" = colors_tmp[3],
  "AUTO_ADJ_Lg" = colors_adj[1], "ES27_ADJ_Lg" = colors_adj[2], "ES64_ADJ_Lg" = colors_adj[3],
  "AUTO_AVG_Lg" = colors_avg[1], "ES27_AVG_Lg" = colors_avg[2], "ES64_AVG_Lg" = colors_avg[3],

  "AUTO_AR_NoLg" = colors_ar[1], "ES27_AR_NoLg" = colors_ar[2], "ES64_AR_NoLg" = colors_ar[3],
  "AUTO_EPI_NoLg" = colors_epi[1], "ES27_EPI_NoLg" = colors_epi[2], "ES64_EPI_NoLg" = colors_epi[3],
  "AUTO_TMP_NoLg" = colors_tmp[1], "ES27_TMP_NoLg" = colors_tmp[2], "ES64_TMP_NoLg" = colors_tmp[3],
  "AUTO_ADJ_NoLg" = colors_adj[1], "ES27_ADJ_NoLg" = colors_adj[2], "ES64_ADJ_NoLg" = colors_adj[3],
  "AUTO_AVG_NoLg" = colors_avg[1], "ES27_AVG_NoLg" = colors_avg[2], "ES64_AVG_NoLg" = colors_avg[3]
)

@@

plot2

# --------- WEEK3 MODELS ------------- #
plot3<-ggplot(W3, aes(x = Best_Result, fill = Best_Result))  +  
  geom_bar() +  
  scale_fill_manual(values = custom_colors) +  
  labs(title = "Best model on the 48 U.S. states (W3)",       
       x = "", y = "Number of states") +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  facet_wrap(~ Model_Type, scales = "free_x")
plot3

@@

 Now we will utilize a Wilcox test with Holm p-value ajustment for repeated comparassion to evaluate the differences in model performances.

1 WEEK AHEAD - Wilcoxon test with Holm p-value adjustment

@@

2 WEEKS AHEAD - Wilcoxon test with Holm p-value adjustment

@@

W4_percentage_of_improvement
```
